---
title: "phenotypes"
author: "steph"
date: "2023-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir="C:/Users/Steph/OneDrive - Texas A&M University/Thrushes/survival/data/")
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



```{r, include=FALSE}

#load packages
pkgs<-c("tidyverse","HIest","viridis","ggpubr","fields","mgcv","car","lme4",
        "MuMIn","psych","qgraph")
sapply(pkgs,library,character.only=T); rm(pkgs)
source("../thrush_survival/00_functions_survival.R")
theme_set(theme_classic())

#read in data
HI_thrush<-read.csv("thrush_survival_230712.csv") #survival data - output from latitude_cleaning.R
colnames(HI_thrush)<-gsub("_",".",colnames(HI_thrush))

phenotypes<-read.csv("thrush_phenotypes_HY_Pemberton.csv")
colnames(phenotypes)<-gsub("_",".",colnames(phenotypes))
phenotypes<-phenotypes%>%
  mutate(ancestry=ancestry*(-1)+1)

phenoNetwork<-read.csv("phenotypeNetwork.csv")


phenoInput<-phenotypes%>%
  select(doy.fall.r1,bearing.fall.1,
         kipps,tail.length,tarsus.length,wing.cord)%>%
  drop_na()

migrateInput<-phenotypes%>%
  select(doy.fall.r1,doy.fall.r2,doy.fall.r4,
         bearing.fall.1,bearing.fall.2,bearing.fall.4,
         kipps,tail.length,tarsus.length,wing.cord)%>%
  drop_na()

poly.all<-data.frame(X.co=rep(c(0,0.5,1)), #make triangle polygon for plotting
                     Y.co=rep(c(0,1,0)))
```

Visualize relationships to ancestry

```{r, eval=FALSE, include=F}

#OLD: size correct linear measurements and normalize everything
#don't do this - we're doing PCAs instead


#get linear trait names
lin.traits<-c("tail.length","wing.cord","kipps","distal","carpal","p9","p10")
tl<-"tarsus.length" #and tarsus length col name

options(na.action = "na.omit")
#size correct linear measurements
for(x in lin.traits){
  phenotypes[,x]<-as.numeric(as.character(phenotypes[,x]))
  beta<-lm(phenotypes[,x]~phenotypes[,tl])$coefficients[2]
  name1<-paste(x,"corr",sep=".")
  phenotypes[,name1]<-phenotypes[,x]-beta*
    (phenotypes[,tl]-mean(phenotypes[,tl],na.rm=T))}


#normalize linear measurements

for(x in c(tl,paste(lin.traits,"corr",sep="."))){ #to run on size corrected traits
  pheno.sub<-phenotypes[!is.na(phenotypes[,x]),c("name.in.vcf",x)]
  pheno.sub[,paste("qnorm",x,sep="_")]<-quantNorm(pheno.sub[,x])
  phenotypes<-left_join(phenotypes,pheno.sub[,c(1,3)],by="name.in.vcf")}

#normalize non-size corrected linear measurements

#for(x in c(tl,paste(lin.traits,"corr",sep="."))){ #to run on size corrected traits
for(x in c(tl,lin.traits)){
  phenotypes[,x]<-as.numeric(as.character(phenotypes[,x]))
  pheno.sub<-phenotypes[!is.na(phenotypes[,x]),c("name_in_vcf",x)]
  pheno.sub[,paste("qnorm",x,sep="_")]<-quantNorm(pheno.sub[,x])
  phenotypes<-left_join(phenotypes,pheno.sub[,c(1,3)],by="name_in_vcf")}


```

```{r, fig.height=8*0.6, fig.width=8}
all.traits=colnames(phenoInput)

ggplot(phenotypes%>%select(name.in.vcf,age.release,ancestry,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
         drop_na(),aes(x=trait.val,fill=as.character(round(ancestry,1))))+
  geom_histogram()+facet_wrap(vars(trait.name),scales="free")+
  scale_fill_viridis(discrete=T,name="ancestry")+
  theme(strip.text.y.left = element_text(angle = 0))

phenoAncestryPlot<-phenotypes%>%select(name.in.vcf,age.release,ancestry,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
  mutate(trait.name=gsub("\\."," ",trait.name))%>%
  mutate(trait.name=gsub("doy","departure day",trait.name))%>%
  mutate(trait.name=gsub("cord","chord",trait.name))%>%
  mutate(trait.name=gsub(" 1","",trait.name))%>%
  mutate(trait.name=gsub(" r1","",trait.name))%>%
         drop_na()

ggplot(phenoAncestryPlot,aes(y=trait.val,x=ancestry))+
  geom_point()+geom_smooth(method="lm",colour="grey20")+
  facet_wrap(vars(trait.name),scales="free")+
  scale_fill_viridis(discrete=T,name="ancestry")+
  theme(strip.text.y.left = element_text(angle = 0))+
  xlab("Trait value")+ylab("Ancestry")

ggplot(phenotypes%>%select(name.in.vcf,age.release,ancestry,sex.binary,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
         drop_na(),aes(y=trait.val,x=sex.binary,colour=as.character(sex.binary)))+
  geom_boxplot(size=1.5)+facet_wrap(vars(trait.name),scales="free")+
  scale_colour_manual(values=c("coral3","cyan4"),name="Sex")+
  theme(strip.text.y.left = element_text(angle = 0))


```

## Mismatch by trait

Leave out adults - no behavioural data for dead adults

```{r, eval=FALSE}


#make adult df 
thrush.adult<-thrush.all%>%filter(age.release=="adult")
mis.adult<-estimate_mismatch(thrush.adult,all.traits,0.1)

adult.pairs<-plot_pairwise_mismatch(mis.adult)
do.call(ggarrange,c(adult.pairs,ncol=4,nrow=5))

```

Then juveniles

```{r, eval=F}

phenotypes<-left_join(phenotypes,HI_thrush%>%select(t5.spring40,name.in.vcf))

mis.juvie<-estimate_mismatch(phenotypes,all.traits,0.1)
mis.juvie.1<-left_join(mis.juvie[[1]],phenotypes)

#plot pairwise mismatch
juvie.pairs<-plot_pairwise_mismatch(mis.juvie)
do.call(ggarrange,c(juvie.pairs,ncol=4,nrow=3))

```

Look at mismatch by survival
Expectation: negative slope

```{r, eval=FALSE}
#plot mismatch by survival

juvie.l<-mis.juvie.1%>%pivot_longer(cols=grep("_|mismatch",colnames(mis.juvie[[1]])),
                                       names_to="mismatch.pairs",values_to="mismatch.vals")

ggplot(data=juvie.l,aes(x=mismatch.vals,y=t5.spring40))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="grey10",formula=y~x,method="glm")+
  facet_wrap(vars(mismatch.pairs))+
  ylab("Survival")+xlab("Mismatch")

out1=data.frame()
for(col1 in grep("_|mismatch",colnames(mis.juvie[[1]]))){
  glm1<-glmer(t5.spring40~mis.juvie.1[,col1]+(1|release.year)+(1|sex.binary),
        data=mis.juvie.1,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out1<-rbind(out1,data.frame(Predictor=colnames(mis.juvie.1)[col1],slope=round(sum1$coefficients[2,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }
out1$a<-ifelse(out1$p.val<0.05,"*","")
out1

```

And traits by survival
Expectation: positive quadratic relationship

```{r,eval=FALSE}

juvie.l.2<-mis.juvie.1%>%pivot_longer(cols=unique(mis.juvie[[2]]$trait),
                           names_to="trait",values_to="trait.vals")
ggplot(data=juvie.l.2,aes(x=trait.vals,y=t5.spring40))+
  geom_smooth(colour="grey10",formula=y~poly(x,2),method="glm")+
  geom_point(colour="grey30",shape=1,stroke=1)+
  facet_wrap(vars(trait))+
  ylab("Survival to breeding grounds")+xlab("Normalized trait value")

out2<-data.frame()
#fit models for each trait
for(col2 in unique(mis.juvie[[2]]$trait)){
  glm1<-glmer(t5.spring40~mis.juvie.1[,col2]+I(mis.juvie.1[,col2]^2)+
                (1|release.year)+(1|sex.binary),
              data=mis.juvie.1,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out2<-rbind(out2,data.frame(Predictor=c(col2,paste(col2,"^2",sep="")),
                              slope=round(sum1$coefficients[2:3,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }
rownames(out2)<-1:nrow(out2)
out2$a<-ifelse(out1$p.val<0.05,"*","")
out2

```

## Summarise phenotype variation along PC axes to account for correlated traits

Select traits with sufficient sample size

```{r}

phenotypes<-left_join(phenotypes,HI_thrush%>%select(name.in.vcf,t5.spring40,t2.fall25))

trait.l<-phenotypes%>%pivot_longer(cols=all_of(all.traits),
                               values_to="trait.val",names_to="trait")%>%
  select(ancestry,heterozygosity,release.year,release.site,t5.spring40,
         t2.fall25,name.in.vcf,trait.val,trait,sex.binary)

keep_traits<-trait.l%>%group_by(trait)%>%filter(!is.na(trait.val))%>%
  summarise(counts=n())%>%filter(counts>50)%>%pull(trait)

trait.l<-trait.l%>%filter(trait%in%keep_traits)
df.pc<-trait.l%>%pivot_wider(names_from=trait,values_from=trait.val)%>%drop_na()
trait.cols<-keep_traits


```

Run principal components analysis

```{r}
#principal components analysis with varimax rotation
pc1<-principal(as.matrix(df.pc[,trait.cols]),nfactors=3,rotate="varimax",scores=T,covar=F)

#pc1$values #eigenvalues - first three <0.9

sum(pc1$values[1:3])/sum(pc1$values)


pc.loadings<-data.frame(matrix(round(pc1$loadings,3),6,3),trait=rownames(pc1$weights))
colnames(pc.loadings)[1:3]<-gsub("R","P",colnames(pc1$weights))

ggplot(pc.loadings%>%pivot_longer(cols=1:3),aes(y=gsub("\\."," ",trait),x=name,fill=abs(value)))+
  geom_tile()+scale_fill_gradient(low="white",high="grey10",name="loading")+
  xlab("PC axis")+ylab("Trait")
```

```{r, fig.width=9,fig.height=9}
biplot(pc1)

```

Look at relationship between PC axes and survival
Expect U-shaped curve (disruptive selection)

```{r,fig.height=2,fig.width=6}

df.pc<-cbind(df.pc,pc1$scores)

axis.names<-c("Wing Shape + Tail","Bearing + Tarsus","Departure Day")

#weird syntax needed because passing variables instead of text
df.pc<-df.pc%>%rename(!!axis.names[1]:=RC1,!!axis.names[2]:=RC2,
               !!axis.names[3]:=RC3)


juvie.l<-df.pc%>%pivot_longer(cols=axis.names,
                              names_to="trait",values_to="trait.vals")


pC<-ggplot(data=juvie.l,aes(x=trait.vals,y=t5.spring40))+
  geom_smooth(colour="grey10",formula=y~x+poly(x,2),method="glm")+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  facet_grid(cols=vars(trait))+
  ylab("Survival")+xlab("PC score")
pC

```

```{r,fig.height=2,fig.width=6}

ggplot(data=juvie.l,aes(x=ancestry,y=trait.vals))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="coral3",method="lm")+
  facet_grid(cols=vars(trait))+
  ylab("PC score")+xlab("Ancestry")


```


And visualize survival to wintering grounds

```{r,fig.height=2,fig.width=6}

pCfall<-ggplot(data=juvie.l,aes(x=trait.vals,y=t2.fall25))+
  geom_smooth(colour="grey10",formula=y~x+poly(x,2),method="glm")+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  facet_grid(cols=vars(trait))+
  ylab("Survival - fall")+xlab("PC score")
pCfall

```

```{r,fig.height=4,fig.width=6,include=F,eval=F}

bearing<-phenotypes%>%pivot_longer(cols=grep("bearing",colnames(phenotypes)),
                              names_to="trait",values_to="trait.vals")

bearing$trait<-gsub("bearing.fall.","latitude ",bearing$trait)

ggplot(data=bearing,aes(x=trait.vals,y=t5.spring40))+
  geom_smooth(colour="grey10",formula=y~x+poly(x,2),method="glm")+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  facet_wrap(vars(trait),scales="free")+
  ylab("Survival")+xlab("Fall bearing")


glm1<-glmer(t5.spring40~phenotypes[,"bearing.fall.5"]+I(phenotypes[,"bearing.fall.5"]^2)+
                (1|release.year)+(1|sex.binary),
              data=phenotypes,family=binomial(link="logit"))
Anova(glm1)
summary(glm1)

```

Test if PC axes determine survival
Fit mixed models with linear and quadratic fixed effects of trait value

```{r}

out4<-data.frame()
for(col1 in axis.names){
  glm1<-glmer(t5.spring40~df.pc[,col1]+I(df.pc[,col1]^2)+
                (1|release.year)+(1|sex.binary),
              data=df.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out4<-rbind(out4,data.frame(Predictor=col1,
                              Predictor2=c("linear","quadratic"),
                              slope=round(sum1$coefficients[2:3,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }
rownames(out4)<-1:nrow(out4)

#write.csv(out4,"../figures/PC_table.csv")
out4



```

Repeat for survival to overwintering sites

```{r}
out4a<-data.frame()
for(col1 in axis.names){
  glm1<-glmer(t2.fall25~df.pc[,col1]+I(df.pc[,col1]^2)+
                (1|release.year)+(1|sex.binary),
              data=df.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out4a<-rbind(out4a,data.frame(Predictor=col1,
                              Predictor2=c("linear","quadratic"),
                              slope=round(sum1$coefficients[2:3,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }
rownames(out4a)<-1:nrow(out4a)

#write.csv(out4a,"../figures/PC_table_fall.csv",row.names=F)
out4a

```

Mismatch between PC axes

```{r, fig.width=4,fig.height=4}


RC.mis<-estimate_mismatch(df.pc,axis.names,0.99)
  
RC.mis.plots<-plot_pairwise_mismatch(RC.mis)
do.call(ggarrange, c(RC.mis.plots))

```

Example mismatch

```{r,fig.height=3,fig.width=6}

ggarrange(RC.mis.plots[[1]],RC.mis.plots[[length(RC.mis.plots)]])

RCdf<-RC.mis[[1]]

parents.w<-RC.mis[[2]]%>%pivot_wider(names_from=trait,values_from=parent.trait.val)

colnames(RCdf)<-gsub("\\+",".",gsub(" ",".",colnames(RCdf)))
colnames(parents.w)<-gsub("\\+",".",gsub(" ",".",colnames(parents.w)))

RCdf$t5.spring40<-as.character(RCdf$t5.spring40)
mis.pair<-"Bearing...Tarsus_Departure.Day"
pB<-ggplot()+
      geom_point(data=RCdf,
                 aes_string(x=str_split(mis.pair,"_")[[1]][1],
                            y=str_split(mis.pair,"_")[[1]][2],
                            colour=mis.pair,shape="t5.spring40"),size=2)+
      geom_point(data=parents.w,aes_string(x=str_split(mis.pair,"_")[[1]][1],
                                           y=str_split(mis.pair,"_")[[1]][2],
                                           shape="pop.type"),size=5,stroke=1.5)+
      scale_shape_manual(values=c(1,19,2,5),name="",guide="none")+
      scale_alpha(range=c(0.3,1))+
      scale_colour_viridis(name="mismatch",option="F")+
      #scale_colour_gradientn(colours=pal)+
      xlab(gsub("\\."," ",gsub("\\.\\.\\."," + ",str_split(mis.pair,"_")[[1]][1])))+
      ylab(gsub("\\."," ",gsub("\\.\\.\\."," + ",str_split(mis.pair,"_")[[1]][2])))+
      theme(aspect.ratio=1)
pB


```

Look at effect of mismatch on survival

```{r,fig.height=4,fig.width=6}

df.pc<-left_join(df.pc,RC.mis[[1]])

#write.csv(df.pc,"phenotype_PCA_mismatch.csv",row.names=F)

juvie.l<-df.pc%>%pivot_longer(cols=grep("_|mismatch",colnames(df.pc)),
                                    names_to="mismatch.pairs",values_to="mismatch.vals")

juvie.l$mismatch.pairs<-gsub("mismatch.all","All Axes",gsub("_"," x\n",juvie.l$mismatch.pairs))

pD<-ggplot(data=juvie.l,aes(x=mismatch.vals,y=t5.spring40))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="grey10",formula=y~x,method="glm")+
  facet_wrap(vars(mismatch.pairs))+
  ylab("Survival")+xlab("Mismatch")

pD

```

And then on fall migration

```{r,fig.height=4,fig.width=6}


pDfall<-ggplot(data=juvie.l,aes(x=mismatch.vals,y=t2.fall25))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="grey10",formula=y~x,method="glm")+
  facet_wrap(vars(mismatch.pairs))+
  ylab("Survival - fall")+xlab("Mismatch")

pDfall

```

And then actually fit survival models

```{r}

out3<-data.frame()
for(col1 in grep("_|mismatch",colnames(df.pc))){
  glm1<-glmer(t5.spring40~df.pc[,col1]+(1|release.year)+(1|sex.binary),
              data=df.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out3<-rbind(out3,data.frame(Predictor=colnames(df.pc)[col1],slope=round(sum1$coefficients[2,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }

out3<-out3%>%
  mutate(Predictor=gsub("_"," X ",Predictor))%>%
  mutate(Predictor=gsub("mismatch.all","Mismatch - all",Predictor))

#write.csv(out3,"../figures/PC_mismatch_table.csv")
out3

```

And on fall migration

```{r}

out3a<-data.frame()
for(col1 in grep("_|mismatch",colnames(df.pc))){
  glm1<-glmer(t2.fall25~df.pc[,col1]+(1|release.year)+(1|sex.binary),
              data=df.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out3a<-rbind(out3a,data.frame(Predictor=colnames(df.pc)[col1],slope=round(sum1$coefficients[2,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }

out3a<-out3a%>%
  mutate(Predictor=gsub("_"," X ",Predictor))%>%
  mutate(Predictor=gsub("mismatch.all","Mismatch - all",Predictor))

#write.csv(out3a,"../figures/PC_mismatch_table_fall.csv",row.names=F)
out3a

```



and make a phenotype network

```{r,eval=F}

phenoNetwork<-phenoNetwork%>%rename("wing\nchord"=cord)

pheno.nodes<-phenoNetwork%>%
  mutate(rowIDs=colnames(phenoNetwork)[1:nrow(phenoNetwork)])%>%
  tibble::column_to_rownames('rowIDs')%>%
  select(1:6)%>%
  mutate_if(is.character, as.numeric)

png(filename="../figures/phenoNetwork.png",width=600,height=400)
qgraph( pheno.nodes,color="white",layout="spring",
           labels=colnames(pheno.nodes),
           minimum=0.1, #minimum edge to plot
           vsize=12, #node size
           label.norm="0000", #normalize node labels to string length
           fade=F,shape="circle",label.scale=T,
           border.color=phenoNetwork$colourPCs,border.width=8,
           negCol="grey60",posCol="grey20",label.color=1,cut=0 )

dev.off()

```



Then fig 3

```{r}


img1<-png::readPNG("../figures/phenoNetwork.png")
pA<-ggplot() + 
    background_image(img1) +
    # This ensures that the image leaves some space at the edges
    #theme(plot.margin = margin(t=1, l=1, r=1, b=1, unit = "cm"))+
  theme(axis.line=element_line(color="white"))

gg1<-ggarrange(pA,pB,ncol=1,nrow=2,heights=c(0.4,0.6),labels=c("A","B"))
gg2<-ggarrange(pC,pD,ncol=1,nrow=2,heights=c(0.4,0.6),labels=c("C","D"))
fig3<-ggarrange(gg1,gg2,widths=c(0.45,0.55))+ bgcolor("white")
fig3
#ggsave(filename="../figures/Fig3.png",device="png",plot=fig3,width=18,height=14,units="cm")
#ggsave(filename="../figures/Fig3.pdf",device="pdf",plot=fig3,width=18,height=14,units="cm")


```

And fig S5

```{r}


gg2fall<-ggarrange(pCfall,pDfall,ncol=1,nrow=2,heights=c(0.4,0.6),labels=c("A","B"))
#ggsave(filename="../figures/FigS5.png",device="png",plot=gg2fall,width=18*0.6,height=14,units="cm")
#ggsave(filename="../figures/Fig3.pdf",device="pdf",plot=fig3,width=18,height=14,units="cm")


```













```{r, include=F}

#clean up variables

rm(aov1,df.pc,glm1,juvie.l,out3,out4,pc.loadings,pc1,RC.mis,RC.mis.plots,sum1,
   trait.l,col1,keep_traits,trait.cols,axis.names)

```


## Again with a smaller sample size and more migratory phenotypes

Select traits with sufficient sample size

```{r}

m.traits<-colnames(migrateInput)

mtrait.l<-phenotypes%>%pivot_longer(cols=all_of(m.traits),
                               values_to="trait.val",names_to="trait")%>%
  select(ancestry,heterozygosity,release.year,release.site,t5.spring40,
         name.in.vcf,trait.val,trait,sex.binary)

keep_traits<-mtrait.l%>%group_by(trait)%>%filter(!is.na(trait.val))%>%
  summarise(counts=n())%>%filter(counts>50)%>%pull(trait)

mtrait.l<-mtrait.l%>%filter(trait%in%keep_traits)
mdf.pc<-mtrait.l%>%pivot_wider(names_from=trait,values_from=trait.val)%>%drop_na()
mtrait.cols<-keep_traits; rm(keep_traits)


```

Run principal components analysis

```{r}
#principal components analysis with varimax rotation
mpc1<-principal(as.matrix(mdf.pc[,mtrait.cols]),nfactors=4,rotate="varimax",scores=T,covar=F)

#mpc1$values #eigenvalues - first four <0.9

mpc.loadings<-data.frame(matrix(round(mpc1$loadings,3),10,4),trait=rownames(mpc1$weights))
colnames(mpc.loadings)[1:4]<-gsub("R","P",colnames(mpc1$weights))

ggplot(mpc.loadings%>%pivot_longer(cols=1:4),aes(y=trait,x=name,fill=abs(value)))+
  geom_tile()+scale_fill_gradient(low="white",high="grey10",name="loading")+
  xlab("PC axis")+ylab("Trait")
```

```{r, fig.width=9,fig.height=9}
biplot(mpc1)

```

Look at relationship between PC axes and survival
Expect U-shaped curve (disruptive selection)

```{r,fig.height=2,fig.width=6}

mdf.pc<-cbind(mdf.pc,mpc1$scores)

m.axis.names<-c("Bearing","Timing","Morphology","Kipps")

#weird syntax needed because passing variables instead of text
mdf.pc<-mdf.pc%>%rename(!!m.axis.names[1]:=RC1,!!m.axis.names[2]:=RC2,
               !!m.axis.names[3]:=RC3,!!m.axis.names[4]:=RC4)


mjuvie.l<-mdf.pc%>%
  pivot_longer(cols=m.axis.names,names_to="trait",values_to="trait.vals")


ggplot(data=mjuvie.l,aes(x=trait.vals,y=t5.spring40))+
  geom_smooth(colour="grey10",formula=y~x+poly(x,2),method="glm")+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  facet_grid(cols=vars(trait))+
  ylab("Survival")+xlab("PC score")

```

Test if PC axes determine survival
Fit mixed models with linear and quadratic fixed effects of trait value

```{r}


out4<-data.frame()
for(col1 in m.axis.names){
  glm1<-glmer(t5.spring40~mdf.pc[,col1]+I(mdf.pc[,col1]^2)+
                (1|release.year)+(1|sex.binary),
              data=mdf.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out4<-rbind(out4,data.frame(Predictor=col1,
                              Predictor2=c("linear","quadratic"),
                              slope=round(sum1$coefficients[2:3,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }
rownames(out4)<-1:nrow(out4)

#out4$a<-ifelse(out4$p.val<0.05,"*","")


#write.csv(out4,"../figures/PC_table.csv")
out4



```

Mismatch between PC axes

```{r}
RC.mis<-estimate_mismatch(mdf.pc,m.axis.names,0.99)
  
RC.mis.plots<-plot_pairwise_mismatch(RC.mis)
do.call(ggarrange, c(RC.mis.plots,ncol=4,nrow=3))

```

Example mismatch

```{r,fig.height=3,fig.width=6}

ggarrange(RC.mis.plots[[1]],RC.mis.plots[[length(RC.mis.plots)]])

```

Look at effect of mismatch on survival

```{r,fig.height=4,fig.width=6}

mdf.pc<-left_join(mdf.pc,RC.mis[[1]])
juvie.l<-mdf.pc%>%pivot_longer(cols=grep("_|mismatch",colnames(mdf.pc)),
                                    names_to="mismatch.pairs",values_to="mismatch.vals")

juvie.l$mismatch.pairs<-gsub("mismatch.all","All Axes",gsub("_"," x\n",juvie.l$mismatch.pairs))

ggplot(data=juvie.l,aes(x=mismatch.vals,y=t5.spring40))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="grey10",formula=y~x,method="glm")+
  facet_wrap(vars(mismatch.pairs))+
  ylab("Survival")+xlab("Mismatch")

```

And then actually fit survival models

```{r}

out3<-data.frame()
for(col1 in grep("_|mismatch",colnames(mdf.pc))){
  glm1<-glmer(t5.spring40~mdf.pc[,col1]+(1|release.year)+(1|sex.binary),
              data=mdf.pc,family=binomial(link="logit"))
  aov1<-Anova(glm1)
  sum1<-summary(glm1)
  out3<-rbind(out3,data.frame(Predictor=colnames(mdf.pc)[col1],slope=round(sum1$coefficients[2,1],2),
                              Chisq=round(aov1$Chisq,2),p.val=round(aov1$`Pr(>Chisq)`,2)))  }

out3$Predictor<-gsub("_"," X ",out3$Predictor)

#write.csv(out3,"../figures/PC_mismatch_table.csv")
out3

```




