---
title: "phenotypes"
author: "steph"
date: "2023-04-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_knit$set(root.dir="C:/Users/Steph/OneDrive - Texas A&M University/Thrushes/survival/data/")
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



```{r, include=FALSE}

#load packages
pkgs<-c("tidyverse","HIest","viridis","ggpubr","fields","mgcv","car","lme4",
        "MuMIn","psych","qgraph")
sapply(pkgs,library,character.only=T); rm(pkgs)
source("../thrush_survival/00_functions_survival.R")
theme_set(theme_classic())

#read in data
HI_thrush<-read.csv("thrush_survival_240106.csv") #survival data - output from survival_cleaning.R
colnames(HI_thrush)<-gsub("_",".",colnames(HI_thrush))
ch_thrush<-read.csv("thrush_survival_ch_240104.csv")

phenotypes<-read.csv("thrush_phenotypes_HY_Pemberton.240104.csv")
colnames(phenotypes)<-gsub("_",".",colnames(phenotypes))
phenotypes<-phenotypes%>%select(-ancestry,-heterozygosity)

phenoInput<-phenotypes%>%
  select(doy.fall.r1,bearing.fall.1,
         kipps,tail.length,tarsus.length,wing.cord)%>%
  drop_na()

migrateInput<-phenotypes%>%
  select(doy.fall.r1,doy.fall.r2,doy.fall.r4,
         bearing.fall.1,bearing.fall.2,bearing.fall.4,
         kipps,tail.length,tarsus.length,wing.cord)%>%
  drop_na()

poly.all<-data.frame(X.co=rep(c(0,0.5,1)), #make triangle polygon for plotting
                     Y.co=rep(c(0,1,0)))


phenotypes<-left_join(phenotypes,HI_thrush%>%
                        select(name.in.vcf,ancestry.no4,heterozygosity,
                              ancestry.scaf4))%>%
  rename(ancestry=ancestry.no4)

```

Visualize relationships of raw phenos to ancestry and sex

```{r, fig.height=8*0.6, fig.width=8}
all.traits=colnames(phenoInput)

ggplot(phenotypes%>%select(name.in.vcf,age.release,ancestry,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
         drop_na(),aes(x=trait.val,fill=as.character(round(ancestry,1))))+
  geom_histogram()+facet_wrap(vars(trait.name),scales="free")+
  scale_fill_viridis(discrete=T,name="ancestry")+
  theme(strip.text.y.left = element_text(angle = 0))

phenoAncestryPlot<-phenotypes%>%select(name.in.vcf,age.release,ancestry,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
  mutate(trait.name=gsub("\\."," ",trait.name))%>%
  mutate(trait.name=gsub("doy","departure day",trait.name))%>%
  mutate(trait.name=gsub("cord","chord",trait.name))%>%
  mutate(trait.name=gsub(" 1","",trait.name))%>%
  mutate(trait.name=gsub(" r1","",trait.name))%>%
         drop_na()

ggplot(phenoAncestryPlot,aes(y=trait.val,x=ancestry))+
  geom_point()+geom_smooth(method="lm",colour="grey20")+
  facet_wrap(vars(trait.name),scales="free")+
  scale_fill_viridis(discrete=T,name="ancestry")+
  theme(strip.text.y.left = element_text(angle = 0))+
  xlab("Trait value")+ylab("Ancestry")

ggplot(phenotypes%>%select(name.in.vcf,age.release,ancestry,sex.binary,all.traits)%>%
         pivot_longer(all.traits,names_to="trait.name",values_to="trait.val")%>%
         drop_na(),aes(y=trait.val,x=sex.binary,colour=as.character(sex.binary)))+
  geom_boxplot(size=1.5)+facet_wrap(vars(trait.name),scales="free")+
  scale_colour_manual(values=c("coral3","cyan4"),name="Sex")+
  theme(strip.text.y.left = element_text(angle = 0))


```

## Summarise phenotype variation along PC axes to account for correlated traits

Select traits with sufficient sample size

```{r}

phenotypes<-left_join(phenotypes,HI_thrush%>%select(name.in.vcf,t5.spring40,t2.fall25))

trait.l<-phenotypes%>%pivot_longer(cols=all_of(all.traits),
                               values_to="trait.val",names_to="trait")%>%
  select(ancestry,heterozygosity,release.year,release.site,t5.spring40,
         t2.fall25,name.in.vcf,trait.val,trait,sex.binary)

keep_traits<-trait.l%>%group_by(trait)%>%filter(!is.na(trait.val))%>%
  summarise(counts=n())%>%filter(counts>50)%>%pull(trait)

trait.l<-trait.l%>%filter(trait%in%keep_traits)
df.pc<-trait.l%>%pivot_wider(names_from=trait,values_from=trait.val)%>%drop_na()
trait.cols<-keep_traits


```

Run principal components analysis

```{r, fig.height=4,fig.width=4}
#principal components analysis with varimax rotation
pc1<-principal(as.matrix(df.pc[,trait.cols]),nfactors=3,rotate="varimax",scores=T,covar=F)

#pc1$values #eigenvalues - first three <1

sum(pc1$values[1:3])/sum(pc1$values)


pc.loadings<-data.frame(matrix(round(pc1$loadings,3),6,3),trait=rownames(pc1$weights))
colnames(pc.loadings)[1:3]<-gsub("R","P",colnames(pc1$weights))

ggplot(pc.loadings%>%pivot_longer(cols=1:3),aes(y=gsub("\\."," ",trait),x=name,fill=abs(value)))+
  geom_tile()+scale_fill_gradient(low="white",high="grey10",name="loading")+
  xlab("PC axis")+ylab("Trait")
```

```{r, fig.width=9,fig.height=9}
biplot(pc1)

df.pc<-cbind(df.pc,pc1$scores)

axis.names<-c("WingShapeTail","BearingTarsus","DepartureDay")

#weird syntax needed because passing variables instead of text
df.pc<-df.pc%>%rename(!!axis.names[1]:=RC1,!!axis.names[2]:=RC2,
               !!axis.names[3]:=RC3)

```

Look at relationship between PC axes and survival
Expect U-shaped curve (disruptive selection)

```{r}

fit_Bear<-read.csv("outputs/hmmCJS.BearingTarsus.fit.phi.20240104.csv")
fit_Day<-read.csv("outputs/hmmCJS.DepartureDay.fit.phi.20240104.csv")
fit_Wing<-read.csv("outputs/hmmCJS.WingShapeTail.fit.phi.20240104.csv")

fit_Bear$trait<-"Bearing + Tarsus"
fit_Day$trait<-"Departure Day"
fit_Wing$trait<-"Wing Shape + Tail"

fit_pheno<-rbind(fit_Bear%>%rename(traitVal=BearingTarsus)%>%
                   select(trait,traitVal,est.adj,lcl,ucl),
      fit_Day%>%rename(traitVal=DepartureDay)%>%
        select(trait,traitVal,est.adj,lcl,ucl),
      fit_Wing%>%rename(traitVal=WingShapeTail)%>%
        select(trait,traitVal,est.adj,lcl,ucl))%>%
  mutate(lcl=lcl^299,ucl=ucl^299)%>%
  mutate(yvar=0)

pC<-ggplot(data=fit_pheno,aes(x=traitVal,y=est.adj))+
  geom_ribbon(aes(ymin=lcl, ymax=ucl),fill="grey80")+
  geom_smooth(colour="grey20",se=0)+
  geom_point(aes(x=traitVal,y=yvar),shape="|",colour="grey20")+
  facet_grid(cols=vars(trait))+ylab("Survival")+xlab("PC Score")

pC  

```

Mismatch between PC axes

```{r, fig.width=8,fig.height=3}
axis.names<-c("WingShapeTail","BearingTarsus","DepartureDay")

RC.mis<-estimate_mismatch(df.pc,axis.names,0.99)
```


```{r, fig.width=4,fig.height=4}  
RC.mis.plots<-plot_pairwise_mismatch(RC.mis)
do.call(ggarrange, c(RC.mis.plots))

```

Example mismatch

```{r,fig.height=3,fig.width=6}

ggarrange(RC.mis.plots[[1]],RC.mis.plots[[length(RC.mis.plots)]])

RCdf<-RC.mis[[1]]

parents.w<-RC.mis[[2]]%>%pivot_wider(names_from=trait,values_from=parent.trait.val)

colnames(RCdf)<-gsub("\\+",".",gsub(" ",".",colnames(RCdf)))
colnames(parents.w)<-gsub("\\+",".",gsub(" ",".",colnames(parents.w)))

RCdf$t5.spring40<-as.character(RCdf$t5.spring40)
mis.pair<-"BearingTarsus_DepartureDay"
pB<-ggplot()+
      geom_point(data=RCdf,
                 aes_string(x=str_split(mis.pair,"_")[[1]][1],
                            y=str_split(mis.pair,"_")[[1]][2],
                            colour=mis.pair),size=2)+
      geom_point(data=parents.w,aes_string(x=str_split(mis.pair,"_")[[1]][1],
                                           y=str_split(mis.pair,"_")[[1]][2],
                                           shape="pop.type"),size=5,stroke=1.5,
                 colour="grey60")+
      scale_shape_manual(values=c(2,5),name="",guide="none")+
      scale_alpha(range=c(0.3,1))+
      scale_colour_viridis(name="mismatch",option="F")+
      #scale_colour_gradientn(colours=pal)+
      xlab(gsub("\\."," ",gsub("\\.\\.\\."," + ",str_split(mis.pair,"_")[[1]][1])))+
      ylab(gsub("\\."," ",gsub("\\.\\.\\."," + ",str_split(mis.pair,"_")[[1]][2])))+
      theme(aspect.ratio=1)
pB


```

Look at effect of mismatch on survival

```{r}

df.pc<-left_join(df.pc,RC.mis[[1]])

#write.csv(df.pc,"phenotype_PCA_mismatch.csv",row.names=F)

ch_thrush<-ch_thrush%>%rename(name.in.vcf=name_in_vcf)%>%
    left_join(df.pc%>%select(name.in.vcf,WingShapeTail,BearingTarsus,DepartureDay,
                             BearingTarsus_DepartureDay,WingShapeTail_BearingTarsus,
                             WingShapeTail_DepartureDay,mismatch.all))%>%
  filter(!is.na(BearingTarsus))%>%select(-X)%>%rename(name_in_vcf=name.in.vcf)
#write.csv(ch_thrush,"thrush_phenotypes_ch_240104.csv")

```


```{r,fig.height=4,fig.width=6}

juvie.l<-df.pc%>%pivot_longer(cols=grep("_|mismatch",colnames(df.pc)),
                                    names_to="mismatch.pairs",values_to="mismatch.vals")

juvie.l$mismatch.pairs<-gsub("mismatch.all","All Axes",gsub("_"," x\n",juvie.l$mismatch.pairs))

pD<-ggplot(data=juvie.l,aes(x=mismatch.vals,y=t5.spring40))+
  geom_jitter(width=0,height=0.1,colour="grey30")+
  geom_smooth(colour="grey10",formula=y~x,method="glm")+
  facet_wrap(vars(mismatch.pairs))+
  ylab("Survival")+xlab("Mismatch")

pD

```
Then fig 3

```{r}


img1<-png::readPNG("C:/Users/Steph/OneDrive - Texas A&M University/Thrushes/survival/figures/Fig3/phenoNetwork.png")
pA<-ggplot() + 
    background_image(img1) +
    # This ensures that the image leaves some space at the edges
    #theme(plot.margin = margin(t=1, l=1, r=1, b=1, unit = "cm"))+
  theme(axis.line=element_line(color="white"))

gg1<-ggarrange(pA,pB,ncol=1,nrow=2,heights=c(0.4,0.6),labels=c("A","B"))
gg2<-ggarrange(pC,pD,ncol=1,nrow=2,heights=c(0.4,0.6),labels=c("C","D"))
fig3<-ggarrange(gg1,gg2,widths=c(0.45,0.55))+ bgcolor("white")
fig3
#ggsave(filename=""C:/Users/Steph/OneDrive - Texas A&M University/Thrushes/survival/figures/Fig3/Fig3.png",device="png",plot=fig3,width=18,height=14,units="cm")
#ggsave(filename="C:/Users/Steph/OneDrive - Texas A&M University/Thrushes/survival/figures/Fig3/Fig3.pdf",device="pdf",plot=fig3,width=18,height=14,units="cm")


```

